\LARGE \bf Stochastic Uncertainty Propagation in Power System Dynamics using Measure-valued Proximal Recursions

### 2022-02-20 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS_Rev1.tex
polishing rebuttal round 1

\subsection{From Anisotropic to Isotropic Degenerate Diffusion}\label{subsec:FromAnisoToIso}
Consider the $2n\times 2n$ matrix
\begin{align}
\bm{\Psi}:= \bm{I}_{2} \otimes \left(\bm{M}\bm{\Sigma}^{-1}\right), \label{DefPsi}	
\end{align}
and define the invertible linear map
\begin{align}
\begin{pmatrix}
\bm{\theta}\\
\bm{\omega}	
\end{pmatrix} \mapsto \begin{pmatrix}
\bm{\xi}\\
\bm{\eta}	
\end{pmatrix} := \bm{\Psi} \begin{pmatrix}
\bm{\theta}\\
\bm{\omega}	
\end{pmatrix}. \label{ChangeOfVar}
\end{align}
Applying It\^{o}'s lemma \cite[Ch. 4.2]{oksendal2013stochastic} to the map \eqref{ChangeOfVar}, and using \eqref{ItoSDEvectorlevel}, we find that the transformed state vector $(\bm{\xi},\bm{\eta})^{\top}$ solves the It\^{o} SDE 
\begin{align}
\begin{pmatrix}
{\mathrm{d}}\bm{\xi}\\
{\mathrm{d}}\bm{\eta}	
\end{pmatrix}
\! = \!\begin{pmatrix}
\bm{\eta}\\
-\nabla_{\bm{\xi}}U(\bm{\xi}) -\nabla_{\bm{\eta}}F\left(\bm{\eta}\right)  	
\end{pmatrix}{\mathrm{d}}t + \!\begin{pmatrix}
 \bm{0}_{n\times n}\\
 \bm{I}_{n}	
 \end{pmatrix}{\mathrm{d}}\bm{w},
\label{XiEtaVectorSDE}	
\end{align}
where the potentials
{\small{\begin{subequations}
\begin{align}
U(\bm{\xi}) &:= \!-\displaystyle\sum_{i=1}^{n}\!\frac{1}{\sigma_{i}}P_{i}\xi_{i} + \!\!\displaystyle\sum_{i,j=1}^{n}\!\!\dfrac{m_{i}}{\sigma_{i}^{2}}k_{ij}\!\left(\!1 \!- \!\cos\!\left(\!\dfrac{\sigma_{i}}{m_{i}}\xi_{i} - \dfrac{\sigma_{j}}{m_{j}}\xi_{j} - \varphi_{ij}\!\!\right)\!\right), \label{defU}\\
F(\bm{\eta}) &:= \dfrac{1}{2}\langle\bm{\eta},\bm{M}^{-1}\bm{\Gamma}\bm{\eta}\rangle. \label{defF}	
\end{align}
\label{NewPotentials}	
\end{subequations}}}

Notice that \eqref{XiEtaVectorSDE} is a mixed conservative-dissipative SDE with \emph{isotropic} degenerate diffusion. In particular, the pushforward of the known initial joint PDF \eqref{FPKInitCond} via $\bm{\Psi}$, is given by
\begin{align}
\tilde{\rho}_{0}(\bm{\xi},\bm{\eta}) := \bm{\Psi}_{\sharp} \rho_{0} 
&= \dfrac{\rho_{0}\left(\bm{\Psi}^{-1} \begin{pmatrix}
\bm{\xi}\\
\bm{\eta}	
\end{pmatrix}\right)}{|\det\left(\bm{\Psi}\right)|} \nonumber\\
&= \dfrac{\rho_{0}\left(\bm{\Sigma}\bm{M}^{-1}\bm{\xi},\bm{\Sigma}\bm{M}^{-1}\bm{\eta}\right)}{\left(\displaystyle\prod_{j=1}^{n}m_{j}/\sigma_{j}\right)^{\!2}},
\label{PushforwardOfIC}	
\end{align}
where we used the standard properties of the Kronecker product. The transient joint state PDF $\tilde{\rho}(t,\bm{\xi},\bm{\eta})$ corresponding to \eqref{XiEtaVectorSDE} solves the PDE initial value problem
\begin{subequations}
\begin{align}
&\dfrac{\partial\tilde{\rho}}{\partial t} = -\langle\bm{\eta},\nabla_{\bm{\xi}}\tilde{\rho}\rangle + \nabla_{\bm{\eta}}\cdot\left(\tilde{\rho}\left(\nabla_{\bm{
\xi}}U(\bm{\xi})+\nabla_{\bm{\eta}}F(\bm{\eta})\right)\right) + \dfrac{1}{2}\Delta_{\bm{\eta}}\tilde{\rho}, \label{NewPDE}\\
&\tilde{\rho}(t=t_{0},\bm{\xi},\bm{\eta}) = \underbrace{\tilde{\rho}_{0}(\bm{\xi},\bm{\eta})}_{\text{from}\;\eqref{PushforwardOfIC}}. \label{NewIC}	
\end{align}
\label{TransformedPDEivp}	
\end{subequations}
In other words, \eqref{TransformedPDEivp} is the macroscopic dynamics corresponding to the sample path dynamics \eqref{XiEtaVectorSDE}.

@@

\subsection{Proximal Update}\label{subsec:ProxUpdate}
Looking at \eqref{NewPotentials} and \eqref{NewPDE}, it is natural to consider an energy functional of the form
\begin{align}
\Phi(\tilde{\rho}) := \!\!\displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\!\!\!\left(\!\bm{M}^{-1}\bm{\Gamma}U(\bm{\xi})+ F\left(\bm{\eta}\right) + \frac{1}{2}\log\tilde{\rho}\!\right)\tilde{\rho}\:\differential\bm{\xi}\:\differential\bm{\eta},
\label{LyapButNotProx}	
\end{align}
which is the sum of a potential energy (expected value of scaled $U$), a weighted kinetic energy (expected value of $F$), and an internal energy (scaled negative entropy, the entropy being $-\int\tilde{\rho}\log\tilde{\rho}$). Indeed, it can be shown that (Appendix \ref{AppLyap}) the functional \eqref{LyapButNotProx} is a Lyapunov-like functional, i.e., $\Phi$ is decreasing along the solution of \eqref{NewPDE}.

However, unlike the gradient drift case mentioned in Sec. \ref{subsec:GenGradDescent}, it is not possible to express the right hand side of \eqref{NewPDE} as the Wasserstein gradient of the functional \eqref{LyapButNotProx}. To see this, recall that the Wasserstein gradient is defined as \cite[Ch. 8]{ambrosio2008gradient} 
$$\nabla^{\text{Wasserstein}}\Phi := -\nabla\cdot\left(\rho\nabla\dfrac{\delta\Phi}{\delta\rho}\right),$$
where $\nabla$ denotes the standard Euclidean gradient w.r.t. the vector $(\bm{\xi},\bm{\eta})^{\top}$, and $\frac{\delta}{\delta\rho}$ denotes the functional derivative. From \eqref{FunctionalDerivative}, \eqref{defvectorfieldv} and \eqref{firstequality} in Appendix A, it is clear that the vector field $\bm{v}$ defined in \eqref{defvectorfieldv} is not equal to the negative of $\nabla\dfrac{\delta\Phi}{\delta\rho}$. Thus, even though the functional \eqref{LyapButNotProx} decreases along the transient solution of \eqref{NewPDE}, we cannot interpret the flow generated by \eqref{TransformedPDEivp} as the Wasserstein gradient flow of the functional \eqref{LyapButNotProx}. Thus, in \eqref{InfiniteDimProxOpDef}, we cannot construct $(\Phi,{\rm{dist}})$ by pairing \eqref{LyapButNotProx} with the Wasserstein metric.

To set up a variational recursion of the form \eqref{InfiniteDimProxOpDef} for \eqref{TransformedPDEivp}, we set $(\Phi,{\rm{dist}}) \equiv \left(\widetilde{\Phi},\widetilde{W}\right)$ where
\begin{align}
\widetilde{\Phi}(\tilde{\rho}) := \int_{\mathbb{T}^{n} \times \mathbb{R}^{n}}\left(F(\boldsymbol{\eta})+\frac{1}{2} \log \tilde{\rho}\right) \tilde{\rho}\: \mathrm{d} \boldsymbol{\xi} \mathrm{d} \boldsymbol{\eta},
\label{Phitilde}	
\end{align}
and 
\begin{align}
\widetilde{W}^{2}\left(\tilde{\varrho}, \tilde{\varrho}_{k-1}\right):=\!&\inf _{\pi \in \Pi\left(\tilde{\varrho}, \tilde{\varrho}_{k-1}\right)}\!\int_{\mathbb{T}^{2 n} \times \mathbb{R}^{2 n}}\!\!\left\{\left\|\overline{\boldsymbol{\eta}}-\boldsymbol{\eta}+h \nabla_{\boldsymbol{\xi}} U(\boldsymbol{\xi})\right\|_{2}^{2}\right.\nonumber\\
&\hspace*{-0.2in}\left.+12\left\|\frac{\overline{\boldsymbol{\xi}}-\boldsymbol{\xi}}{h}-\frac{\overline{\boldsymbol{\eta}}+\boldsymbol{\eta}}{2}\right\|_{2}^{2}\right\} \mathrm{d} \pi(\boldsymbol{\xi}, \boldsymbol{\eta}, \overline{\boldsymbol{\xi}}, \overline{\boldsymbol{\eta}}).
\label{Wtilde}	
\end{align}
In \eqref{Wtilde}, $\Pi\left(\tilde{\varrho}, \tilde{\varrho}_{k-1}\right)$ denotes the collection of all joint PDFs supported on $\mathbb{T}^{2 n} \times \mathbb{R}^{2 n}$ that have finite second raw moments, with the first marginal $\tilde{\varrho}$, and the second marginal $\tilde{\varrho}_{k-1}$.

That the sequence of functions $\{\tilde{\varrho}_{k}\}$ for $k\in\mathbb{N}$, generated by the proximal recursion
\begin{align}
\tilde{\varrho}_{k} &= {\text{prox}}^{\widetilde{W}}_{h\widetilde{\Phi}}(\tilde{\varrho}_{k-1}) \nonumber\\
&\equiv \underset{\tilde{\rho}\in\mathcal{P}_{2}}{\arg\inf}\:\frac{1}{2}\widetilde{W}^{2}\left(\tilde{\varrho},\tilde{\varrho}_{k-1}\right) + h\:\widetilde{\Phi}(\tilde{\rho}), \quad \tilde{\varrho}_{0} := \tilde{\rho}_{0},
\label{KineticProx}	
\end{align}
converges to the flow generated by \eqref{TransformedPDEivp}, i.e., 
$$\tilde{\varrho}_{k}(\bm{\xi},\bm{\eta}) \xrightarrow{h\downarrow 0} {\blue{\tilde{\rho}}}(t=kh,\bm{\xi},\bm{\eta})\;\text{in}\;L^{1}\left(\mathbb{T}^{n}\times\mathbb{R}^{n}\right),$$
was established in \cite{duong2014conservative}. To numerically perform the recursion \eqref{KineticProx}, we employ the proximal algorithm proposed in \cite{caluya2019TAC} with finite number of samples, as explained next.

@@

\appendices
\section{Showing $\Phi(\tilde{\rho})$ is Decreasing along the Flow Generated by \eqref{TransformedPDEivp}}\label{AppLyap}
Let $\tilde{\rho}(t,\bm{\xi},\bm{\eta})$ be the transient solution of \eqref{TransformedPDEivp}. Define
$$\bm{\ell}_{\bm{\xi}} := \frac{1}{2}\nabla_{\bm{\xi}}\log\tilde{\rho}, \quad \bm{\ell}_{\bm{\eta}} := \frac{1}{2}\nabla_{\bm{\eta}}\log\tilde{\rho},$$
and rewrite \eqref{NewPDE} as
$$\dfrac{\partial\tilde{\rho}}{\partial t} = - \nabla\cdot\left(\tilde{\rho}\bm{v}\right)$$
where $\nabla$ is the $2n\times 1$ Euclidean gradient w.r.t. the vector $(\bm{\xi},\bm{\eta})^{\top}$, and the vector field
\begin{align}
\bm{v} := \begin{pmatrix}
 \bm{\eta}\\
 -\nabla_{\bm{\xi}}U(\bm{\xi}) - \bm{M}^{-1}\bm{\Gamma}\bm{\eta} - \bm{\ell}_{\bm{\eta}}	
 \end{pmatrix}.
\label{defvectorfieldv}	
\end{align}

Now consider the time derivative of the functional \eqref{LyapButNotProx} along \eqref{NewPDE}. Following \cite[Proposition 1]{krichene2020global}, 
\begin{align}
\frac{\differential}{\differential t}\Phi = \displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}} \bigg\langle\nabla\dfrac{\delta\Phi}{\delta\tilde{\rho}}, \tilde{\rho}\bm{v}\bigg\rangle\:\differential\bm{\xi}\differential\bm{\eta},
\label{dPhidt}	
\end{align}
wherein the functional derivative
\begin{align}
\dfrac{\delta\Phi}{\delta\tilde{\rho}}	= \!\bm{M}^{-1}\bm{\Gamma}U(\bm{\xi})+ F\left(\bm{\eta}\right) + \frac{1}{2} + \frac{1}{2}\log\tilde{\rho}.
\label{FunctionalDerivative}
\end{align}
Recalling the explicit form of $F(\bm{\eta})$ from \eqref{defF}, and using \eqref{FunctionalDerivative}, we simplify \eqref{dPhidt} as
\begin{strip}
\hrule
\begin{subequations}
\begin{align}
\frac{\differential}{\differential t}\Phi &= \!\displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\!\bigg\langle\begin{pmatrix}
\bm{M}^{-1}\bm{\Gamma}\nabla_{\bm{\xi}}U(\bm{\xi}) + \bm{\ell}_{\bm{\xi}}\\
\bm{M}^{-1}\bm{\Gamma}\bm{\eta} + \bm{\ell}_{\bm{\eta}}	
\end{pmatrix}, \begin{pmatrix}
 \bm{\eta}\\
 -\nabla_{\bm{\xi}}U(\bm{\xi}) - \bm{M}^{-1}\bm{\Gamma}\bm{\eta} - \bm{\ell}_{\bm{\eta}}	
 \end{pmatrix}\bigg\rangle\tilde{\rho}(t,\bm{\xi},\bm{\eta})\:\differential\bm{\xi}\:\differential\bm{\eta}	\label{firstequality}\\
&= \!\displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\!\bigg\{-\|\bm{M}^{-1}\bm{\Gamma}\bm{\eta}\|_{2}^{2} - \|\bm{\ell}_{\bm{\eta}}\|_{2}^{2} - 2\langle \bm{M}^{-1}\bm{\Gamma}\bm{\eta},\bm{\ell}_{\bm{\eta}}\rangle + \langle \bm{\ell}_{\bm{\xi}},\bm{\eta}\rangle - \langle\bm{\ell}_{\bm{\eta}},\nabla_{\bm{\xi}}U(\bm{\xi})\rangle\bigg\}\tilde{\rho}(t,\bm{\xi},\bm{\eta})\:\differential\bm{\xi}\:\differential\bm{\eta} \label{lastbutoneequality}\\
&= -\displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\|\bm{M}^{-1}\bm{\Gamma}\bm{\eta} + \bm{\ell}_{\bm{\eta}}\|_{2}^{2}\:\tilde{\rho}(t,\bm{\xi},\bm{\eta})\:\differential\bm{\xi} \:\differential\bm{\eta}, \label{lastequality}
\end{align}
\label{ExplicitdPhidt}	
\end{subequations}
\hrule	
\end{strip}
where \eqref{lastequality} follows from the fact that the integrals involving $\langle \bm{\ell}_{\bm{\xi}},\bm{\eta}\rangle$ and $\langle\bm{\ell}_{\bm{\eta}},\nabla_{\bm{\xi}}U(\bm{\xi})\rangle$ in \eqref{lastbutoneequality} are zero, see \cite[Appendix D.3]{krichene2020global}.

>>>>

\subsection{From Anisotropic to Isotropic Degenerate Diffusion}\label{subsec:FromAnisoToIso}
Consider the $2n\times 2n$ matrix
\begin{align}
\bm{\Psi}:= \bm{I}_{2} \otimes \left(\bm{M}\bm{\Sigma}^{-1}\right), \label{DefPsi}	
\end{align}
and define the invertible linear map
\begin{align}
\begin{pmatrix}
\bm{\theta}\\
\bm{\omega}	
\end{pmatrix} \mapsto \begin{pmatrix}
\bm{\xi}\\
\bm{\eta}	
\end{pmatrix} := \bm{\Psi} \begin{pmatrix}
\bm{\theta}\\
\bm{\omega}	
\end{pmatrix}. \label{ChangeOfVar}
\end{align}
Applying It\^{o}'s lemma \cite[Ch. 4.2]{oksendal2013stochastic} to the map \eqref{ChangeOfVar}, and using \eqref{ItoSDEvectorlevel}, we find that the transformed state vector $(\bm{\xi},\bm{\eta})^{\top}$ solves the It\^{o} SDE 
\begin{align}
\begin{pmatrix}
{\mathrm{d}}\bm{\xi}\\
{\mathrm{d}}\bm{\eta}	
\end{pmatrix}
\! = \!\begin{pmatrix}\!
\bm{\eta}\\
-\bm{\Upsilon}\nabla_{\bm{\xi}}U(\bm{\xi}) -\nabla_{\bm{\eta}}F\left(\bm{\eta}\right)  	
\!\end{pmatrix}\!{\mathrm{d}}t + \!\begin{pmatrix}\!
 \bm{0}_{n\times n}\\
 \bm{I}_{n}	
 \!\end{pmatrix}\!{\mathrm{d}}\bm{w},
\label{XiEtaVectorSDE}	
\end{align}
where {\blue{the diagonal matrix $\bm{\Upsilon}:= \left(\prod_{i=1}^{n}\sigma_{i}^{2}m_{i}^{-2}\right)\bm{M\Sigma}^{-2}$ has positive entries along the main diagonal, and}} the potentials
{\small{\begin{subequations}
\begin{align}
U(\bm{\xi}) &:= \left(\displaystyle\sum_{i<j}k_{ij}\left(\!1 \!- \!\cos\!\left(\!\dfrac{\sigma_{i}}{m_{i}}\xi_{i} - \dfrac{\sigma_{j}}{m_{j}}\xi_{j} - \varphi_{ij}\!\!\right)\!\right) \right.\nonumber\\
&\left.\qquad\qquad-\displaystyle\sum_{i=1}^{n}\!\frac{\sigma_{i}}{m_{i}}P_{i}\xi_{i}\right)\left(\prod_{i=1}^{n}\frac{m_{i}^{2}}{\sigma_{i}^{2}}\right), \label{defU}
%\!-\displaystyle\sum_{i=1}^{n}\!\frac{1}{\sigma_{i}}P_{i}\xi_{i} + \!\!\displaystyle\sum_{i,j=1}^{n}\!\!\dfrac{m_{i}}{\sigma_{i}^{2}}k_{ij}\!\left(\!1 \!- \!\cos\!\left(\!\dfrac{\sigma_{i}}{m_{i}}\xi_{i} - \dfrac{\sigma_{j}}{m_{j}}\xi_{j} - \varphi_{ij}\!\!\right)\!\right)
\\
F(\bm{\eta}) &:= \dfrac{1}{2}\langle\bm{\eta},\bm{\Sigma}^{-1}\bm{\Gamma}\bm{\eta}\rangle. \label{defF}	
\end{align}
\label{NewPotentials}	
\end{subequations}}}

Notice that \eqref{XiEtaVectorSDE} is a mixed conservative-dissipative SDE with \emph{isotropic} degenerate diffusion. In particular, the pushforward of the known initial joint PDF \eqref{FPKInitCond} via $\bm{\Psi}$, is given by
\begin{align}
\tilde{\rho}_{0}(\bm{\xi},\bm{\eta}) := \bm{\Psi}_{\sharp} \rho_{0} 
&= \dfrac{\rho_{0}\left(\bm{\Psi}^{-1} \begin{pmatrix}
\bm{\xi}\\
\bm{\eta}	
\end{pmatrix}\right)}{|\det\left(\bm{\Psi}\right)|} \nonumber\\
&= \dfrac{\rho_{0}\left(\bm{\Sigma}\bm{M}^{-1}\bm{\xi},\bm{\Sigma}\bm{M}^{-1}\bm{\eta}\right)}{\left(\displaystyle\prod_{i=1}^{n}m_{i}/\sigma_{i}\right)^{\!2}},
\label{PushforwardOfIC}	
\end{align}
where we used the standard properties of the Kronecker product. The transient joint state PDF $\tilde{\rho}(t,\bm{\xi},\bm{\eta})$ corresponding to \eqref{XiEtaVectorSDE} solves the PDE initial value problem
\begin{subequations}
\begin{align}
&\dfrac{\partial\tilde{\rho}}{\partial t} = -\langle\bm{\eta},\nabla_{\bm{\xi}}\tilde{\rho}\rangle + \nabla_{\bm{\eta}}\cdot\left(\tilde{\rho}\left(\bm{\Upsilon}\nabla_{\bm{
\xi}}U(\bm{\xi})+\nabla_{\bm{\eta}}F(\bm{\eta})\right)\right) \nonumber\\
&\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad+ \dfrac{1}{2}\Delta_{\bm{\eta}}\tilde{\rho}, \label{NewPDE}\\
&\tilde{\rho}(t=t_{0},\bm{\xi},\bm{\eta}) = \underbrace{\tilde{\rho}_{0}(\bm{\xi},\bm{\eta})}_{\text{from}\;\eqref{PushforwardOfIC}}. \label{NewIC}	
\end{align}
\label{TransformedPDEivp}	
\end{subequations}
In other words, \eqref{TransformedPDEivp} is the macroscopic dynamics corresponding to the sample path dynamics \eqref{XiEtaVectorSDE}.

@@

\subsection{Proximal Update}\label{subsec:ProxUpdate}
Looking at \eqref{NewPotentials} and \eqref{NewPDE}, it is natural to consider an energy functional of the form
\begin{align}
\Phi(\tilde{\rho}) &:= \!\!\displaystyle\int_{{\blue{\left(\prod_{i=1}^{n}[0,2\pi m_{i}/\sigma_{i})\right)}}\times\mathbb{R}^{n}}\!\!\!\left(\!U(\bm{\xi})+ F\left(\bm{\eta}\right) + \frac{1}{2}\log\tilde{\rho}\!\right)\nonumber\\
&\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\tilde{\rho}\:\differential\bm{\xi}\:\differential\bm{\eta},
\label{LyapButNotProx}	
\end{align}
which is the sum of a potential energy (expected value of $U$), a weighted kinetic energy (expected value of $F$), and an internal energy (scaled negative entropy, the entropy being $-\int\tilde{\rho}\log\tilde{\rho}$). %Indeed, it can be shown that (Appendix \ref{AppLyap}) the functional \eqref{LyapButNotProx} is a Lyapunov-like functional, i.e., $\Phi$ is decreasing along the solution of \eqref{NewPDE}.
However, unlike the gradient drift case mentioned in Sec. \ref{subsec:GenGradDescent}, it is not possible to express the right hand side of \eqref{NewPDE} as the Wasserstein gradient of the functional \eqref{LyapButNotProx}. This can be verified via direct computation by recalling that the Wasserstein gradient is defined as \cite[Ch. 8]{ambrosio2008gradient} 
$$\nabla^{\text{Wasserstein}}\Phi := -\nabla\cdot\left(\rho\nabla\dfrac{\delta\Phi}{\delta\rho}\right),$$
where $\nabla$ denotes the standard Euclidean gradient w.r.t. the vector $(\bm{\xi},\bm{\eta})^{\top}$, and $\frac{\delta}{\delta\rho}$ denotes the functional derivative. %From \eqref{FunctionalDerivative}, \eqref{defvectorfieldv} and \eqref{firstequality} in Appendix A, it is clear that the vector field $\bm{v}$ defined in \eqref{defvectorfieldv} is not equal to the negative of $\nabla\dfrac{\delta\Phi}{\delta\rho}$. 
Thus, 
%even though the functional \eqref{LyapButNotProx} decreases along the transient solution of \eqref{NewPDE}, 
we cannot interpret the flow generated by \eqref{TransformedPDEivp} as the Wasserstein gradient flow of the functional \eqref{LyapButNotProx}. Consequently, in \eqref{InfiniteDimProxOpDef}, we cannot construct $(\Phi,{\rm{dist}})$ by pairing \eqref{LyapButNotProx} with the Wasserstein metric.

To set up a variational recursion of the form \eqref{InfiniteDimProxOpDef} for \eqref{TransformedPDEivp}, we set $(\Phi,{\rm{dist}}) \equiv \left(\widetilde{\Phi},\widetilde{W}_{h,\bm{\Upsilon}}\right)$ where
\begin{align}
\widetilde{\Phi}(\tilde{\rho}) := \!\!\int_{{\blue{\left(\prod_{i=1}^{n}[0,2\pi m_{i}/\sigma_{i})\right)}} \times \mathbb{R}^{n}}\!\!\left(\!F(\boldsymbol{\eta})+\frac{1}{2} \log \tilde{\rho}\!\right) \tilde{\rho}\: \mathrm{d} \boldsymbol{\xi} \mathrm{d} \boldsymbol{\eta},
\label{Phitilde}	
\end{align}
and 
\begin{align}
&\widetilde{W}_{h,\bm{\Upsilon}}^{2}\left(\tilde{\varrho},\tilde{\varrho}_{k-1}\right) := \!\!\underset{\pi\in\Pi\left(\tilde{\varrho},\tilde{\varrho}_{k-1}\right)}{\inf}\!\displaystyle\int_{{\blue{\left(\prod_{i=1}^{n}[0,2\pi m_{i}/\sigma_{i})\right)}} \times \mathbb{R}^{2n}}\!\!s_{h,\bm{\Upsilon}}\left(\bm{\xi},\bm{\eta},\bar{\bm{\xi}},\bar{\bm{\eta}}\right)\nonumber\\
&\differential\pi\left(\bm{\xi},\bm{\eta},\bar{\bm{\xi}},\bar{\bm{\eta}}\right),
\label{Wtilde}    
\end{align}
where $h>0$ is the step-size in proximal recursion, and $\Pi\left(\tilde{\varrho},\tilde{\varrho}_{k-1}\right)$ denotes the set of joint probability measures over the product space $\mathbb{T}^{2n} \times \mathbb{R}^{2n}$ that have finite second moments  with the first marginal $\tilde{\varrho}$, and the second marginal $\tilde{\varrho}_{k-1}$. The ``ground cost" $s_{h,\bm{\Upsilon}}$ in \eqref{Wtilde} is given by
 \begin{align}
     &s_{h,\bm{\Upsilon}}\left(\bm{\xi},\bm{\eta},\bar{\bm{\xi}},\bar{\bm{\eta}}\right):=\nonumber\\
    &\bigg\langle\!\!\left(\bar{\bm{\eta}}-\bm{\eta}+h\bm{\Upsilon} \nabla_{\bm{\xi}} \bm{U}(\bm{\xi})\right),\bm{\Upsilon}^{-1}\!\left(\bar{\bm{\eta}}-\bm{\eta}+h\bm{\Upsilon} \nabla_{\bm{\xi}}\bm{U}(\bm{\xi})\right)\!\!\bigg\rangle\nonumber\\
    & + 12 \bigg\langle\!\left(\!\frac{\bar{\bm{\xi}}-\bm{\xi}}{h}-\frac{\bar{\bm{\eta}}-\bm{\eta}}{h}\!\right),\bm{\Upsilon}^{-1}\!\left(\!\frac{\bar{\bm{\xi}}-\bm{\xi}}{h}-\frac{\bar{\bm{\eta}}-\bm{\eta}}{h}\!\right)\!\bigg\rangle.
    \end{align}

%$\Pi\left(\tilde{\varrho}, \tilde{\varrho}_{k-1}\right)$ denotes the collection of all joint PDFs supported on $\mathbb{T}^{2 n} \times \mathbb{R}^{2 n}$ that have finite second raw moments,

That the sequence of functions $\{\tilde{\varrho}_{k}\}$ for $k\in\mathbb{N}$, generated by the proximal recursion
\begin{align}
\tilde{\varrho}_{k} &= {\text{prox}}^{\widetilde{W}}_{h\widetilde{\Phi}}(\tilde{\varrho}_{k-1}) \nonumber\\
&\equiv \underset{\tilde{\varrho}\in\mathcal{P}_{2}}{\arg\inf}\:\frac{1}{2}\widetilde{W}^{2}\left(\tilde{\varrho},\tilde{\varrho}_{k-1}\right) + h\:\widetilde{\Phi}(\tilde{\varrho}), \quad \tilde{\varrho}_{0} := \tilde{\rho}_{0},
\label{KineticProx}	
\end{align}
converges to the flow generated by \eqref{TransformedPDEivp}, i.e., 
$$\tilde{\varrho}_{k}(\bm{\xi},\bm{\eta}) \xrightarrow{h\downarrow 0} {\blue{\tilde{\rho}}}(t=kh,\bm{\xi},\bm{\eta})\;\text{in}\;L^{1}\left(\mathbb{T}^{n}\times\mathbb{R}^{n}\right),$$
was established in \cite{duong2014conservative}. To numerically perform the recursion \eqref{KineticProx}, we employ the proximal algorithm proposed in \cite{caluya2019TAC} with finite number of samples, as explained next.

@@

%\appendices
%\section{Showing $\Phi(\tilde{\rho})$ is Decreasing along the Flow Generated by \eqref{TransformedPDEivp}}\label{AppLyap}
%Let $\tilde{\rho}(t,\bm{\xi},\bm{\eta})$ be the transient solution of \eqref{TransformedPDEivp}. Define
%$$\bm{\ell}_{\bm{\xi}} := \frac{1}{2}\nabla_{\bm{\xi}}\log\tilde{\rho}, \quad \bm{\ell}_{\bm{\eta}} := \frac{1}{2}\nabla_{\bm{\eta}}\log\tilde{\rho},$$
%and rewrite \eqref{NewPDE} as
%$$\dfrac{\partial\tilde{\rho}}{\partial t} = - \nabla\cdot\left(\tilde{\rho}\bm{v}\right)$$
%where $\nabla$ is the $2n\times 1$ Euclidean gradient w.r.t. the vector $(\bm{\xi},\bm{\eta})^{\top}$, and the vector field
%\begin{align}
%\bm{v} := \begin{pmatrix}
% \bm{\eta}\\
% -\nabla_{\bm{\xi}}U(\bm{\xi}) - \bm{M}^{-1}\bm{\Gamma}\bm{\eta} - \bm{\ell}_{\bm{\eta}}	
% \end{pmatrix}.
%\label{defvectorfieldv}	
%\end{align}
%
%
%Now consider the time derivative of the functional \eqref{LyapButNotProx} along \eqref{NewPDE}. Following \cite[Proposition 1]{krichene2020global}, 
%\begin{align}
%\frac{\differential}{\differential t}\Phi = \displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}} \bigg\langle\nabla\dfrac{\delta\Phi}{\delta\tilde{\rho}}, \tilde{\rho}\bm{v}\bigg\rangle\:\differential\bm{\xi}\differential\bm{\eta},
%\label{dPhidt}	
%\end{align}
%wherein the functional derivative
%\begin{align}
%\dfrac{\delta\Phi}{\delta\tilde{\rho}}	= \!\bm{M}^{-1}\bm{\Gamma}U(\bm{\xi})+ F\left(\bm{\eta}\right) + \frac{1}{2} + \frac{1}{2}\log\tilde{\rho}.
%\label{FunctionalDerivative}
%\end{align}
%Recalling the explicit form of $F(\bm{\eta})$ from \eqref{defF}, and using \eqref{FunctionalDerivative}, we simplify \eqref{dPhidt} as
%\begin{strip}
%\hrule
%\begin{subequations}
%\begin{align}
%\frac{\differential}{\differential t}\Phi &= \!\displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\!\bigg\langle\begin{pmatrix}
%\bm{M}^{-1}\bm{\Gamma}\nabla_{\bm{\xi}}U(\bm{\xi}) + \bm{\ell}_{\bm{\xi}}\\
%\bm{M}^{-1}\bm{\Gamma}\bm{\eta} + \bm{\ell}_{\bm{\eta}}	
%\end{pmatrix}, \begin{pmatrix}
% \bm{\eta}\\
% -\nabla_{\bm{\xi}}U(\bm{\xi}) - \bm{M}^{-1}\bm{\Gamma}\bm{\eta} - \bm{\ell}_{\bm{\eta}}	
% \end{pmatrix}\bigg\rangle\tilde{\rho}(t,\bm{\xi},\bm{\eta})\:\differential\bm{\xi}\:\differential\bm{\eta}	\label{firstequality}\\
%&= \!\displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\!\bigg\{-\|\bm{M}^{-1}\bm{\Gamma}\bm{\eta}\|_{2}^{2} - \|\bm{\ell}_{\bm{\eta}}\|_{2}^{2} - 2\langle \bm{M}^{-1}\bm{\Gamma}\bm{\eta},\bm{\ell}_{\bm{\eta}}\rangle + \langle \bm{\ell}_{\bm{\xi}},\bm{\eta}\rangle - \langle\bm{\ell}_{\bm{\eta}},\nabla_{\bm{\xi}}U(\bm{\xi})\rangle\bigg\}\tilde{\rho}(t,\bm{\xi},\bm{\eta})\:\differential\bm{\xi}\:\differential\bm{\eta} \label{lastbutoneequality}\\
%&= -\displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\|\bm{M}^{-1}\bm{\Gamma}\bm{\eta} + \bm{\ell}_{\bm{\eta}}\|_{2}^{2}\:\tilde{\rho}(t,\bm{\xi},\bm{\eta})\:\differential\bm{\xi} \:\differential\bm{\eta}, \label{lastequality}
%\end{align}
%\label{ExplicitdPhidt}	
%\end{subequations}
%\hrule	
%\end{strip}
%where \eqref{lastequality} follows from the fact that the integrals involving $\langle \bm{\ell}_{\bm{\xi}},\bm{\eta}\rangle$ and $\langle\bm{\ell}_{\bm{\eta}},\nabla_{\bm{\xi}}U(\bm{\xi})\rangle$ in \eqref{lastbutoneequality} are zero, see \cite[Appendix D.3]{krichene2020global}.
%
%From \eqref{ExplicitdPhidt}, it follows that $\frac{\differential}{\differential t}\Phi \leq 0$, i.e., $\Phi$ is non-increasing along the flow generated by \eqref{TransformedPDEivp}. Furthermore, \eqref{lastequality} tells us that $\frac{\differential}{\differential t}\Phi = 0$ if and only if 
%$$\bm{M}^{-1}\bm{\Gamma}\bm{\eta} + \bm{\ell}_{\bm{\eta}} = 0,$$
%i.e., $\Phi$ is stationary if and only if $\tilde{\rho}$ has $\bm{\eta}$ marginal $\int\tilde{\rho}\:\differential\bm{\xi} \propto \exp\left(-\bm{\eta}^{\top}\bm{M}^{-1}\bm{\Gamma}\bm{\eta}\right)$. Following the steps as in \cite[Appendix E.1]{krichene2020global}, this is achieved at the stationary solution of \eqref{NewPDE} with $\bm{\xi}$ marginal $\int\tilde{\rho}\:\differential\bm{\eta} \propto \exp\left(-2U(\bm{\xi})\right)$. In other words, the stationary solution of \eqref{NewPDE} must be of the form 
